---
output: md_document
---

```{r}
knitr::opts_chunk$set(comment='.')
```



# dmutate
Mutate a `data.frame`, adding random variates.

```{r, message=FALSE}
library(dplyr)
library(dmutate)
```


### Univariate examples

Some variables to use in formulae:
```{r, message=FALSE}
low_wt <- 70
high_wt <- 90
mu_wt <- 80
sd <- 60
p.female <- 0.24
```

Use `mutate_random` to implement formulae in data frame.  We can put bounds on any simulated variable
```{r,message=FALSE}
data.frame(ID=1:10) %>% 
  mutate_random(WT[low_wt,high_wt] ~ rnorm(mu_wt,sd))
```

We can simulate from any probability distirbution in `R`
```{r,message=FALSE}
data.frame(ID=1:10) %>% mutate_random(X ~ rcauchy(0,0.5))
```

We can add the variate at any level
```{r,message=FALSE}
data.frame(ID=1:10) %>%
  mutate(GROUP = ID%%2) %>%
  mutate_random(STUDY_RE ~ rnorm(50,sqrt(50))|GROUP)
```

### Simulate multivariate normal with bounds
```{r}
mu <- c(2,200)
Sigma <- diag(c(10,1000))
XY <- X[0,] + Y[200,300] ~ rmvnorm(mu,Sigma)
```
The object
```{r}
XY
```


Simulate
```{r}
data.frame(ID=1:10000) %>%
  mutate_random(XY) %>% 
  summary
```



### An extended example
```{r, message=FALSE}
data.frame(ID=1:10) %>%
  mutate(GROUP = ID%%2) %>%
  mutate_random(WT[low_wt,high_wt] ~ rnorm(mu_wt,1)) %>%
  mutate_random(STUDY_RE ~ rnorm(0,sqrt(50))|GROUP) %>%
  mutate_random(SEX ~ rbinomial(p.female)) %>%
  mutate_random(sigma ~ rgamma(1,1)) %>%
  mutate_random(kappa ~ rgamma(1,1)|GROUP) %>% signif(3)

```


# Create formulae with `expr` to calculate new columns in the `data.frame` using `dplyr::mutate`

We can easily save formulae to `R` variables.  We collect formulae together into sets called `covset`.  For better control for 
where objects are found, we can specify an environment where objects can be found.



```{r}
a <- X ~ rnorm(50,3)
b <- Y ~ expr(X/2 + c)
d <- A+B ~ rlmvnorm(log(c(20,80)),diag(c(0.2,0.2)))
cov1 <- covset(a,b,d)
e <- list(c=3)
```

Notice that `b` has function `expr`.  This assigns the column named `Y` (in this case) to the result of evaluating the 
expression in the data frame using `dplyr::dmutate`.

```{r}
.data <- data.frame(ID=1:3)

mutate_random(.data,cov1,envir=e) %>% signif(3)
```





